---
title: "AugspurgerSynchrony"
author: "sagesteppe"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
fig_caption: yes
vignette: >
  %\VignetteIndexEntry{AugspurgerSynchrony}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package includes a single function which serves to calculate Augspurgers Index of Synchrony.

The package is installable from github. 

```{r Install package, eval = F}
# install.packages('devtools', dependencies = T)
devtools::install_github('sagesteppe/AugspurgerIndex') # one option 
# if for some reason installing devtools to your device is challenging, 
# remotes also offers an install_github function
# install.packages('remotes', dependencies = T)
# remotes::install_github('sagesteppe/AugspurgerIndex') 
```

Once installed the package can be loaded just as any other would. 
```{r attach package to session, message = F, warning = F}
library(AugspurgerIndex)
```

The data should be in Day of year (DOY) format. 
You can obtain values like this using lubridate::ydoy. As we will show below, this function can also convert back to a standard date format. 
Please note that lubridate `yday` starts counting from January 1st, at '1'. 
Also please note because their are leap years, it is always best practice to include the year in the date.  

```{r}
lubridate::yday( as.Date('2011-01-01', '%Y-%m-%d') )  # jan 1st, first day of year
lubridate::yday( as.Date('2011-12-31', '%Y-%m-%d') )  # dec 31st, first day of year
lubridate::yday( as.Date('2011-10-11', '%Y-%m-%d') ) # october 6th, 284 day of year
```


```{r import and view structure of data}
data(flowering_data, package = "AugspurgerIndex")
head(flowering_data[,1:4])
```

```{r run function}
synchrony <- augs_synchrony(dataset = flowering_data,
                             frst_day = flower_start, lst_day = flower_end,
                             year_samp = year, species )
synchrony[12:30,c(1,8:9)]
```

You may need to convert your values to back to calendar date using something like lubridate::yday

```{r}

synchrony |>
  dplyr::group_by(species) |> 
  dplyr::mutate(ID = 1:dplyr::n(),
                flower_start_DOY = lubridate::yday(flower_start),
                flower_end_DOY   = lubridate::yday(flower_end)) |>
  head()
  
```

If you use this function you can cite is using the details below. 
An easy name for the bibtex lookup `@` is benkendorf2024augspurger. 

```{r}
citation('AugspurgerIndex')
```

